---
title: "Tylers.Activity.MarkDown"
output: html_document
---

## Load and transform the data
```{r load.transform}
setwd("/Users/tylersummers/Desktop/Coursera/R") ##set wd
act<-read.csv("activity.csv", header=TRUE) ##assume in working         directory
act$date<-as.Date(act$date,"%Y-%m-%d") ##transform to class = date
```

## Histogram of the total steps per day  

```{r day.hist}
x1<-tapply(act$steps, act$date, sum, na.rm=TRUE)
    ##create array of total steps/day

hist(x1, main="Total Steps by Day from October - November 2012", xlab="Daily Total Steps") 
    ##plot histogram

rug(x1) ##show density within histogram

m<-ceiling(mean(x1)) ##calculate the mean, rounded

m2<-median(x1) ## calculate the median, rounded
```
The mean steps in a day is `r m` and the median is `r m2` steps per day.  
  
##Average daily step pattern  
```{r step.pattern}
library(ggplot2)

y1<-tapply(act$steps, act$interval, mean, na.rm=TRUE)
    ##average steps by intervals across all days

plot(names(y1),y1, type="n", main="Average Interval Steps", 
     xlab= "Intervals", ylab="Average Steps")
    ##set up plot for timeseries 
lines(names(y1),y1) ##add the data

y2<-sort(y1, decreasing=FALSE) ##order interval average steps

y3<-names(y2[length(y2)]) ##interval of max avg steps
````
On average, the most steps occur at the `r y3` interval.  
  
## Missing values imputed  

````{r missing1}
miss<-c(mean(is.na(act$steps)),mean(is.na(act$date)),
        mean(is.na(act$interval))) ##calculate missing in %

z1<-rbind(names(act),miss) ##show results in a df
````
Our data has missing values. See the table of missing values as percentages below.  
`r z1`  
  
##Imputed Data
```` {r missing2}
library(plyr)
impute<-function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
    ##function to replace NAs with mean where mean ignores NAs
act2<-ddply(act, ~interval, transform, steps=impute(steps))
    ##use impute func across intervals and replace in df
a1<-tapply(act2$steps, act2$date, sum, na.rm=TRUE)
    ##create new array of total steps/day

hist(x1, main="Total Steps by Day from October - November 2012 (Imputed)", xlab="Daily Total Steps") 
    ##plot histogram

a2<-ceiling(mean(a1)) ##compute new mean, rounded
a3<-ceiling(median(a1)) ##compute new median, rounded
````
The mean steps in a day is `r a2` and the median is `r a3` steps per day (imputed).  

## Weekdays verses Weekends  
````{r days}
wkdys<-c("Monday","Tuesday","Wednesday","Thursday","Friday")
    ##vector of weekday names
act2$week<-factor((weekdays(act$date) %in% wkdys), 
                  levels= c(FALSE, TRUE), 
                  labels= c("weekend","weekday"))
    ##create column denoting day type

par(mfrow=c(1,2)) ##set view for side by side plots

##plot 1 for weekdays
act3<-subset(act2, act2$week == "weekday") ##weekday subset

w1<-tapply(act3$steps, act3$interval, mean, na.rm=TRUE)
    ##average steps by intervals across all weekdays
plot(names(w1),w1, type="n", main=" Weekday Average Steps", 
     xlab= "Intervals", ylab="Average Steps")
    ##set up plot for timeseries 
lines(names(w1),w1) ##add the data

act4<-subset(act2, act2$week == "weekend") ##weekend subset

w2<-tapply(act4$steps, act4$interval, mean, na.rm=TRUE)
    ##average steps by intervals across weekends
plot(names(w2),w2, type="n", main="Weekend Average Steps", 
     xlab= "Intervals", ylab="Average Steps")
    ##set up plot for timeseries 
lines(names(w2),w2) ##add the data
````
